%Copyright (c) 2015 Alan Chandler, all rights reserved
%
%This file is part of PASv5, an implementation of the Patient Administration
%System used to support Accuvision's Laser Eye Clinics.
%
%PASv5 is licenced to Accuvision (and its successors in interest) free of royality payments
%and in perpetuity in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the
%implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE. Accuvision
%may modify, or employ an outside party to modify, any of the software provided that
%this modified software is only used as part of Accuvision's internal business processes.
%
%The software may be run on either Accuvision's own computers or on external computing
%facilities provided by a third party, provided that the software remains soley for use
%by Accuvision (or by potential or existing customers in interacting with Accuvision).

\documentclass[Draft]{akc}
\author{Alan Chandler}
\title{Project Manual}
\project{PI Audio Recorder}
\date{23 February 2020}
\usepackage{listings}
\lstdefinestyle{bstyle}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=L,
  columns=flexible,
  xleftmargin=\parindent,
  language=bash,
  showstringspaces=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
}
\lstset{style=bstyle}
\begin{document}
\maketitle
\abstract{%
<<<<<<< HEAD
PAS v5 is going to run on a Raspberry Pi Model B (Version 2 or later). This document describes
the hardward that is needed, the set up of the operating system on that hardware, And how to load PAS v5 software on to it and make it operational
=======
This is a project to create a portable audio recorder.  It consists of a Raspberry PI 4 running a basic server system and a nodejs application which 
provides a web interface to control recording and to spawn the ffmpeg application to record the audio, but also generate a video stream of the volume
on the incoming audio. Recorded audio files will be recoverable via sftp for offline post processing. 
>>>>>>> 3eaa5e83709238fdaef8522ec372b15fa2671746
}
\tableofcontents
\section{Introduction}

<<<<<<< HEAD
I have decided that PASv5 should run on a standalone Rapsberry Pi, There are a number of reasons forsudo
this decision, but mainly because the main server already runs a web server that is incompatible
with PASv5 and because the manager of the main server has asked that we run PASv5 on a separate
machine,

Because of the low cost of a Raspberry PI server (the proposed configuration costs less that £70
inclusive of everything needed) and because the only data it will store is logs to aid in the
debugging of PASv5 the server itself does not have to be kept alive with expensive support.  It
would be perfectly feasible, indeed this is what is happening, for Accuvision to keep a spare
machine in the cupboard fully configured and ready to go should the main one fail.  This can prevent
any issues of putting PASv5 into production.

The purpose of this document is to describe how to create this server from scratch should it ever be
necessary, initially setting up the hardware and a system image which is loaded on to an SD card.
Secondly we then show how to set the infrastructure that enables PASv5 to be built, run in
production, and updated to the next release when available.

From time to time will we use the system tools on the server to update the software that is not part
of PASv5. When we do it is then necessary to take a backup of this revised system image so that can
be installed on backup hardware.  We describe how to do this.

=======
>>>>>>> 3eaa5e83709238fdaef8522ec372b15fa2671746
\section{The Hardware}

The hardware needed for the PASv5 Server is fairly minimal and is given in the list below:-

\begin{itemize}
\item Raspberry Pi 2 Model B - Or Pi 3
\item FLIRC Raspberry Pi B+ Case (recommended for its thermal properties, although any case will work)
\item Raspberry Pi Universal Power Supply (5v, 2A, microUSB)
\item 16GB Micro SD (SanDisk class @10 recommended)
\item Ethernet cable to connect to the internal lan
\item (Optional) 64B USB stick for backups.
\end{itemize}

One version of this is needed for the main server, it is recommeded to get a second copy (excluding
the optional usb stick) to act as an immediate replacement in the event of failure.

\subsection{Putting an Operating System and other Software on the Hardware}

The hardware boots and runs the software installed on a MicroSD card placed in the slot in the case
for that purpose.  The first partition of the MicroSD card \emph{must} be of type FAT32 and contain
the boot image and information associated with booting the system.  The other partitions make up the
running system.

You start with a blank card and add information to it.

In essence there are two ways to create the SD card
\begin{enumerate}
\item From scratch, creating the partitions, installing an operating system image that has been
downloaded from the internet, and then installing all the software that is required.  This
\emph{will} require another machine.  Dependant on the peripherals (SD Card Reader or not) on the
`preparation' machine a USB to SD card converter  \emph{might} might be needed.
\item Copying a backup image taken previously on to the new SD Card.  Assuming this process is done
on the live system a USB SD card converter \emph{will} be required.
\end{enumerate}

These two options are described in the subsections below.

\section{Creating an SD Card}
\subsection{Creating an Original SD Card Operating System Image}
\subsubsection{Getting The Initial Skeleton Sysem onto the SD card}
Working on the `preparation' machine, download a working image of the Rasbian (Linux, Debian) System.  At the time of writing
this is a version called `Buster' and is downloadable from
\url{https://www.raspberrypi.org/downloads/raspbian/}.

This page shows various downloads, but the `Raspian Buster Lite' package is the one required. This starts
with a minimal number of packages installed (more can be added later). Chose the zip file version,
which if you click on the link will load
\href{https://downloads.raspberrypi.org/raspbian_lite_latest} {The latest Zip File Version}.

When this zip file has finished downloading extract the contents (an `.img' file).
To install this image on the card we need to follow the instructions given here:-

\begin{lstlisting}
sudo dd status=progress bs=4M if=imagefilename.img of=/dev/sdX  #X = whatever the device holding the sd card is
\end{lstlisting}

This will create an SD card with two partitions filling up about 1.4GB of the 16GB on the card.
Before the this card can be put in the Raspberry Pi space needs to be allocated for one more
partition.  The easiest way to do that is with  \texttt{fdisk}.  Ultimately the second partition is
going to take the remainder of the SD card minus a 1GB partition at the end used to hold the swap
file\footnote{We are going to be using the `btrfs' filesystem that will not work with swap. It is
therefore necessary to create an swap partition}.  It is necessary to
take the scary step of deleting the second partition and then recreating it at exactly the same
start location, but with the end location 1GB (we try to stay on 4M boundaries in order to line files up
with the erase boundaries of the sd card) from the end of the card. Provided we do recreate this
second partition at exactly the same place as the original, no data will be lost from it.

Follow the following steps:-
\begin{enumerate}
\item With the card still in the /dev/sdX as above, run \texttt{fdisk /dev/sdX}.
\item Press `p' to print the partition list and note down the starting sector of partition 2.  The
list should also give sector number for the last sector on the card. Subtract 2097152\footnote{1GB
in 512B sectors} from it and round this down so it exactly divisible by 8192. This as the start sector
for partition 3.
\item Press `d' to delete this partition.
\item Press `n' to recreate this partition.  Make it a primary partiion and use the starting sector
noted in step 2 above and the end sector as the value also noted in Step 2 minus 1. Don't remove the
ext4 signature when prompted.
\item Press `p' to re print the partition list and note the ending sector number for the second partition.
\item Press `n' to add a new partition starting at one sector after the ending sector noted above.
Again it should be a primary partition and should fill the rest of the disk.
\item Press `t' to specify the type of this partition as `82'. which is the type of a swap partition.
\item Press `w' to save the partition table back to the card and exit the program.
\end{enumerate}

We can now turn partition 3 into a swap partiion
\begin{lstlisting}
sudo mkswap -L swap /dev/sdX3
\end{lstlisting}

We should now have an image that will at least run in a Raspberry Pi

\subsubsection{Initial Adjustments}

Put the card into the Raspberry Pi, borrow and connect a keyboard and monitor to it and power it on.  This
should boot up and present a log on prompt. Log on to the ``pi'' account using the password ``raspberry''.

Once logged on the following has to be completed. This can be done with the command \texttt{sudo
raspi-config}:-

\begin{itemize}
\item Update \texttt{raspi-config} to the latest version (Advanced Options/Update)
\item Change the password of the ``pi'' account. I have used the main
``Administrative Logon'' password given in document ``Passwords for the SQL Server Implementation of
PAS'' written by me in April 2012 and lodged with Accuvision at the time.
\item Set boot options to Console, requiring user to log in.
\item Ensure the international settings are for English(British).
\item Disable all interface options except SSH server, which should be enabled.
\item Give the unit a hostname of ``pas'' (Advanced Options/Hostname).
\item Set the required GPU memory to 16  (Advanced Options/Memory Split)
\end{itemize}

Exit the \texttt{raspi-config} program and tell the Raspberry Pi to reboot itself.  From this point
on direct connection of the keyboard and monitor is no longer needed\footnote{Unless an error is
made that has to be recovered from.} and they can be disconnected and returned to where ever they
were borrowed from!

Now use SSH to connect to the ``pi'' account.  This will be something like \texttt{ssh pi@pas}\footnote{If
the host computer has been used on other sd cards connected to this same host then it may be necessary to
edit ~/ssh/known\_hosts on the host computer to remove old and invalid host keys}.

The following set of tasks need to be performed

\subsubsection{Check that the entire root partition is filled with a filesystem}

Raspbian seems to do this automatically on first start, but it is worth double checking by
issuing the following command:-

\begin{lstlisting}
sudo resize2fs /dev/mmcblk0p2
\end{lstlisting}

This will either do the expansion or report there is nothing to do.
\subsubsection{Use the swap partition}
We can now get the system to use the swap partition with the following steps:-
\begin{lstlisting}
sudo systemctl stop dphys-swapfile
sudo systemctl disable dphys-swapfile
sudo nano /etc/fstab          #Add the line shown below to the file and remove the comment about no swap because of dphys-swapfile
swapon -a
# if all works
sudo apt remove dphys-swapfile
sudo apt autoremove
sudo rm /var/swap
\end{lstlisting}

The line to add the \texttt{/etc/fstab} should look like this
\begin{lstlisting}
LABEL=swap      none        swap  sw                0    0
\end{lstlisting}


\subsection{Convert to btrfs}
\subsubsection{An Overview of the Process}
Previously the process of creating a btrfs filesystem from the contents of the existing ext4 root filesystem was
automatic using the \texttt{btrfs-convert} utility.  However that is no longer supplied and is not recommended either,
so we have to take a new approach.  So we take this in a series of steps
\begin{enumerate}
  \item prepare an \texttt{initramfs} image that includes the btrfs filesystem
  \item adjust the parameters to enable the new system to correctly boot
  \item move offline so we can manipulate the root filesystem
  \item backup the existing ext4 rootfilesystem, delete it and replace it with a btrfs one
  \item make btrfs subvolumes for root and log
  \item restore the rootfilesystem into the new btrfs subvolumes
  \item replace it in the raspberry pi and boot it.
\end{enumerate}

\subsubsection{Setting up initramfs to include the btrfs filesystem module}

The first step is to install the btrfs and initramfs tools that we need. It is also helpful to
install vim, the text editor that we will use and rsync.

\begin{lstlisting}
  sudo apt install vim btrfs-tools initramfs-tools
\end{lstlisting}

Edit \texttt{/etc/initramfs-tools/modules} to add btrfs as the last line. Then use

\begin{lstlisting}
sudo mkinitramfs -o /boot/initramfs7.img $(uname -r)
\end{lstlisting}

To create an initramfs that can be used to start up with btrfs

Edit \texttt{/boot/config.txt} file to have the following line at the top

\begin{lstlisting}
initramfs initramfs7.img followkernel
\end{lstlisting}

Edit \texttt{/boot/cmdline.txt} to add the the following two elements \texttt{rootfstype=btrfs} and
\texttt{rootflags=subvol=\_root}

It is also important that whenever the kernel is upgraded we recreate our \texttt{initramfs7.img}
file for the newer kernel.

Create a new file \texttt{/etc/kernel/postinst.d/initramfs-rebuild} with the following contents and
then make it executable.

\begin{lstlisting}
  #!/bin/sh -e

  # Rebuild initramfs{7,7l,8).img after kernel upgrade to include new kernel's modules.
  # Exit if rebuild cannot be performed
  [ -x /usr/sbin/mkinitramfs ] || exit 0

  # Exit if not building kernel for this Raspberry Pi's hardware version (assume armv7).
  version="$1"
  case "${version}" in
    *-v7+) arch=7 ;;
    *-v7l+) arch=7l ;;
    *-v8+) arch=8 ;;
    *) exit 0
  esac
  # Exit if rebuild not needed.
  [ -f /boot/initramfs$arch.img ] && lsinitramfs /boot/initramfs$arch.img | grep -q "/$version$" && exit 0  # Already in initramfs.

  # Rebuild.
  mkinitramfs -o /boot/initramfs$arch.img "$version"
\end{lstlisting}

Make the file executable.

Shutdown the system and remove the SD card.
\subsubsection{Backup existing ext4 filesystem and replace with new btrfs filesysem}

The following steps need to take place without the SD card as the running kernel. As they are purely
file system changes.  It can be done either on the original host machine where the SD card was
formated in the first place, or on the Raspberry Pi using another SD card made from the original
image and used as a host.

Assume that there is a spare mount point at \texttt{/mnt} in whatever environment is being used and that the
SD card we are converting is in \texttt{/dev/sdX}, then the
following commands should be followed

\begin{lstlisting}
sudo -s   #transition to root mode and stay there
mount /dev/sdX2 /mnt
cd ~
mkdir rootfs-backup
rsync -axAHX /mnt/ rootfs-backup/  #backup root
umount /dev/sdX2
wipefs -a /dev/sdX2
mkfs -t btrfs -L rootfs /dev/sdX2
mount -o subvol=/ /dev/sdX2 /mnt
btrfs subvolume create /mnt/_root
btrfs subvolume create /mnt/_log
btrfs quota enable /mnt/_log              #so we can apply a quota to it
btrfs qgroup limit 2g /mnt/_log           #of 2 Gigs
btrfs subvolume create /mnt/_pas          #to be used when we come to set up pas user
rsync -axAHX rootfs-backup/ /mnt/_root/   #restore filesystem
mv /mnt/_root/var/log/*  /mnt/_log/       #separate logs
mkdir -p /mnt/_root/mnt/btrfs_pool        #create a place to mount the whole btrfs filesystem
mkdir -p /mnt/_root/mnt/backup            #create a directory for mounting external backup media in future
vim /mnt/_root/etc/fstab  #set the new contents of /etc/fstab as listed below
umount /mnt
\end{lstlisting}

The \texttt{/etc/fstab} file edited above should look like the following

\begin{lstlisting}
proc            /proc     proc  defaults                          0    0
LABEL=boot      /boot     vfat  defaults                          0    2
LABEL=rootfs    /         btrfs defaults,noatime,ssd,subvol=_root 0    0
LABEL=rootfs    /var/log	btrfs defaults,noatime,ssd,subvol=_log  0    0
LABEL=swap      none      swap  sw                                0    0
\end{lstlisting}

Remove the SD card and place in back in the Raspberry Pi and boot it.  If this has worked and you can ssh into the pi
account, you can also safely delete the rootfs-backup directory created earlier


\subsection{Adjusting the system so it ready for PAS}

\subsubsection{Set up SSH Keys to avoid having to enter a password on each connection}

It is desirable to be able to connect as the ``pi'' user to perform system maintenance tasks without having
to enter a password every time.  This can be done securely by providing the public keys of users
allowed to connect and perform this function. At the time of writing this document these are:-

\begin{itemize}
\item Alan from his home machine
\item Alan from the Windows 10 Virtual Machine, enabling access to the clinic over an encrypted network.
\item Alan from the machine in the Clinic
\end{itemize}

Each of these accounts are configured already with private keys to connect, so the public keys need to be copied
across and installed in an \texttt{authorized\_keys} file.

Firstly whilst in the ``pi'' account perform the following commands:-

\begin{lstlisting}
mkdir .ssh
chmod 700 .ssh
\end{lstlisting}

Then back on the machine where all the keys are stored \footnote{The `pas' without any extension is the private key for this box and can be used to connect from this box to elsewhere. This will be used, for instance to connect to a master copy of the PASv5 repository and pull in the latest versions.}

\begin{lstlisting}
scp kanga.pub roo.pub accwrk7.pub pas pi@pas:.ssh/
\end{lstlisting}

Then ssh back to pi@pas

\begin{lstlisting}
cd .ssh
mv pas id_rsa  #set up the private key for this unit
cat kanga.pub roo.pub accwrk7.pub > authorized_keys
\end{lstlisting}

It should now be possible to connect directly to pi@pas without entering a password\footnote{This
assumes that the private keys corresponding the the copied public keys are without pass phrases.
This is generally the case with the keys used just because of the convenience.}.

\subsubsection{Add additional software needed to run the system}

Initially upgrade to latest versions of software

\begin{lstlisting}
sudo apt-get update
sudo apt-get upgrade
sudo reboot
\end{lstlisting}

In the clinic, the default domain being used is `accuvision.local'.  Because this clashes with the
default local domain (`.local') it can cause some strange results when looking up addresses. To
solve this we need to edit the \texttt{/etc/nsswitch.conf} file and change the `hosts` line to
the following:-

\begin{lstlisting}
hosts:          files dns mdns4_minimal [NOTFOUND=return]
\end{lstlisting}

Then install the additional software required

\begin{lstlisting}
  sudo apt-get install git htop wget postfix
  \end{lstlisting}

The only one of these that needs any configuration is postfix, and the following files
should be set up as follows:-

\texttt{/etc/postfix/main.cf}
\begin{lstlisting}
  # See http://www.postfix.org/COMPATIBILITY_README.html -- default to 2 on
  # fresh installs.
  compatibility_level = 2

  smtpd_banner = $myhostname ESMTP $mail_name (Raspbian)
  biff = no
  # appending .domain is the MUA's job.
  append_dot_mydomain = no

  delay_warning_time = 4h
  confirm_delay_cleared = yes
  notify_classes = delay
  readme_directory = no

  # TLS parameters
  smtpd_tls_cert_file=/etc/ssl/certs/ssl-cert-snakeoil.pem
  smtpd_tls_key_file=/etc/ssl/private/ssl-cert-snakeoil.key
  smtpd_use_tls=yes
  smtpd_tls_session_cache_database = btree:${data_directory}/smtpd_scache
  smtp_tls_session_cache_database = btree:${data_directory}/smtp_scache

  smtp_use_tls = yes
  smtp_sasl_auth_enable = yes
  smtp_sasl_password_maps = hash:/etc/postfix/relay_password
  smtp_tls_session_cache_database = btree:${data_directory}/smtp_scache
  smtp_sasl_security_options =
  smtp_tls_security_level = may

  smtpd_relay_restrictions = permit_mynetworks permit_sasl_authenticated defer_unauth_destination
  myhostname = pas.accuvision.local
  alias_maps = hash:/etc/aliases
  alias_database = hash:/etc/aliases
  myorigin = /etc/mailname

  relayhost = [mail.accuvision.co.uk]
  mynetworks = 127.0.0.0/8 [::ffff:127.0.0.0]/104 [::1]/128
  mailbox_size_limit = 0
  recipient_delimiter = +
  inet_interfaces = loopback-only
  inet_protocols = all

  #There are no local domains - everything must go somewhere else, the virtual maps define it
  local_transport = error: User unknown
  local_recipient_maps =
  mydestination =

  #virtual alias domains, we are the mail server for the domain, but everything gets forwarded
  virtual_alias_domains =
    $myhostname

  virtual_alias_maps = hash:/etc/postfix/virtual
\end{lstlisting}

\texttt{/etc/postfix/virtual}

\begin{lstlisting}
  #Local transport will specify virtual, so these entries reroute
  mailer-daemon alan.chandler+pas@hartley-consultants.com
  postmaster  alan.chandler+pas@hartley-consultants.com
  webmaster  alan.chandler+pas@hartley-consultants.com
  abuse  alan.chandler+pas@hartley-consultants.com
  root alan.chandler+pas@hartley-consultants.com
  pas alan.chandler+pas@hartley-consultants.com
  pi alan.chandler+pas@hartley-consultants.com

\end{lstlisting}

\texttt{/etc/postfix/relay\_password}

\begin{lstlisting}
  mail.accuvision.co.uk info@accuvision.co.uk:EyECare555
\end{lstlisting}

It is also necessary to run a few commands
\begin{lstlisting}
  sudo hostname pas.accuvision.local
  sudo postmap /etc/postfix/virtual
  sudo postmap /etc/postfix/relay_password
\end{lstlisting}

\subsubsection{Create the pas user}

We are going to have a user ``pas'' who will own and run the pas system.  However, we want to give this
user their own btrfs subvolume because it will be easier to back this up and restore it.  However in order to get the permissions
set up correctly we have create the user first and then move all the file so a new subvolume
\begin{lstlisting}
  sudo -s   #become root
  adduser pas
  mount -o subvol=_pas /dev/mmcblk0p2 /mnt
  mv /home/pas/* /mnt/
  mv /home/pas/.* /mnt/
  umount /mnt
  mount -o subvol=_pas /dev/mmcblk0p2 /home/pas
  vim /etc/fstab     #add the line below into the exsting file
\end{lstlisting}

The following should be added to /etc/fstab
\begin{lstlisting}
  LABEL=rootfs  /home/pas	      btrfs   defaults,noatime,ssd,subvol=_pas   0       0
\end{lstlisting}

\subsubsection{Set up backup processes}

We want to create some cron jobs that help maintain the system.  The first of these will be a
filesystem check.  Create a file called \texttt{/etc/cron.monthly/btrfsmaintenance} with the
following content:-

\begin{lstlisting}
#!/bin/sh

btrfs scrub start -Bq /dev/mmcblk0p2

\end{lstlisting}

To sit alongside this file we also need to call \texttt{fstrim}.  However with the buster release of
the operating system we can do this simply by
\begin{lstlisting}
sudo systemctl enable --now fstrim.timer
\end{lstlisting}

Now create a file called \texttt{/etc/cron.weekly/backup} with the following content:-

\begin{lstlisting}
  #!/bin/bash
  mount /dev/mmcblk0p2 -o subvol=/ /mnt/btrfs_pool

  if [ -d /mnt/btrfs_pool/backup-pas-week5 ]; then
    btrfs subvolume delete /mnt/btrfs_pool/backup-pas-week5 2>&1 > /dev/null
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-week4 ]; then
    mv /mnt/btrfs_pool/backup-pas-week4 /mnt/btrfs_pool/backup-pas-week5
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-week3 ]; then
    mv /mnt/btrfs_pool/backup-pas-week3 /mnt/btrfs_pool/backup-pas-week4
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-week2 ]; then
    mv /mnt/btrfs_pool/backup-pas-week2 /mnt/btrfs_pool/backup-pas-week3
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-week1 ]; then
    mv /mnt/btrfs_pool/backup-pas-week1 /mnt/btrfs_pool/backup-pas-week2
  fi

  btrfs subvolume snapshot -r /mnt/btrfs_pool/_pas /mnt/btrfs_pool/backup-pas-week1 > /dev/null
  umount /mnt/btrfs_pool

\end{lstlisting}

and a file called \texttt{/etc/cron.month/backup} with the following content:-

\begin{lstlisting}
  #!/bin/bash
  mount /dev/mmcblk0p2 -o subvol=/ /mnt/btrfs_pool

  if [ -d /mnt/btrfs_pool/backup-rootfs-month3 ]; then
    btrfs subvolume delete -c /mnt/btrfs_pool/backup-rootfs-month3 2>&1 > /dev/null
  fi
  if [ -d /mnt/btrfs_pool/backup-rootfs-month2 ]; then
    mv /mnt/btrfs_pool/backuk-rootfs-month2 /mnt/btrfs_pool/backup-rootfs-month3
  fi
  if [ -d /mnt/btrfs_pool/backup-rootfs-month1 ]; then
    mv /mnt/btrfs_pool/backup-rootfs-month1 /mnt/btrfs_pool/backup-rootfs-month2
  fi
  btrfs subvolume snapshot -r /mnt/btrfs_pool/_root /mnt/btrfs_pool/backup-rootfs-month1 > /dev/null

  if [ -d /mnt/btrfs_pool/backup-pas-month3 ]; then
    btrfs subvolume delete -c /mnt/btrfs_pool/backup-pas-month3  2>&1 > /dev/null
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-month2 ]; then
    mv /mnt/btrfs_pool/backuk-pas-month2 /mnt/btrfs_pool/backup-pas-month3
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-month1 ]; then
    mv /mnt/btrfs_pool/backup-pas-month1 /mnt/btrfs_pool/backup-pas-month2
  fi
  if [ -d /mnt/btrfs_pool/backup-pas-week5 ]; then
    btrfs subvolume snapshot -r /mnt/btrfs_pool/backup-pas-week5 /mnt/btrfs_pool/backup-pas-month1
  fi

  umount /mnt/btrfs_pool

\end{lstlisting}

Make sure all three files have excute priviledges


\subsubsection{Prepare for setting up and running production system}

We need to be able to easily connect via ssh to the pas user.  Set this up with
the following commands:-

\begin{lstlisting}
sudo -s
cd /home/pas
mkdir .ssh
chown pas.pas .ssh
chmod 700 .ssh
cd .ssh
cp /home/pi/.ssh/* .
chown pas.pas *
\end{lstlisting}

Now connect as the ``pas'' user (\texttt{ssh pas@pas}), firstly to check that it works, but also
to perform the next few commands to setup ``nvm'' a node version manager and use it to install the latest
(or a new) version of `node` along with some key `node' based command line tools.

Perform the following commands:-

\begin{lstlisting}
wget -qO- https://raw.githubusercontent.com/creationix/nvm/v0.33.0/install.sh | bash
vim .bashrc #add the following line to the end of the file
export PAS_ENVIRO=clinic
# logout and back in again
nvm install v10.16.3       #use the latest version in which pas has been tested
npm install -g pm2 eslint eslint-config eslint-plugin-html babel-eslint mocha polymer-cli
#logout and login as pi
sudo setcap 'cap_net_bind_service=+ep' /home/pas/.nvm/versions/node/v10.16.3/bin/node  #allow node to bind to low ports
\end{lstlisting}

We now need to clone pasv5 from the master repository, and set it up for production.  ssh in again as the pas user.

\begin{lstlisting}
git clone pas@asgard.hartley-consultants.com:pasv5.git
cd pasv5cd
vim .git/hooks/post-commit  #See contents below
chmod +x .git/hooks/post-commit
ln -s -T .git/hooks/post-commit .git/hooks/post-merge
git checkout production
npm install
polymer build
vi .git/hooks/post-commit #Remove the comment # for the three lines related to oldbuild
pm2 start pas.json
crontab -e  #setup evening task to run every day
pm2 install pm2-logrotate
pm2 set pm2-logrotate:max_size 1M
pm2 set pm2-logrotate:retain 5
pm2 set pm2-logrotate:rotateInterval '15 1 1 1 * *'
pm2 startup #note output from this command and run it as user pi (who has sudo capability)
pm2 save #needed because fails when running as user pi
\end{lstlisting}

The following should be the contents of the post-commit hook created above

\begin{lstlisting}
#!/bin/bash
#
#  For production branch, build the application

branch=$(git branch | sed  -n s/^\*\ //p)
version=$(git describe --tags)

cd "$(git rev-parse --show-cdup)"
npm install

if [ "$branch" == "production" ]; then
    npm install
    # first adjust client code to run from previous version whilst we build the new
#    npm run enviro:old
#    mv build oldbuild
    #build the new
    polymer build
    #swap over to it
    npm run enviro
    #remove old client
#    rm -rf oldbuild
else
    npm run enviro
fi
\end{lstlisting}

The following should be the contents of the crontab edited above.

\begin{lstlisting}
NODE_EXEC = /home/pas/.nvm/versions/node/v10.16.3/bin/node
EVENING = /home/pas/pasv5/server/evening.js
WEBENQ = /home/pas/pasv5/server/getwebenqs.js
MAILTO = alan.chandler@hartley-consultants.com

0 19 * * * $NODE_EXEC $EVENING
*/2 8-18 * * 1-6 $NODE_EXEC $WEBENQ
\end{lstlisting}
\section{Updating Node to a New Version}

\begin{lstlisting}
nvm install <new-version>
nvm alias default <new-version>
\end{lstlisting}
Logon as pi user
\begin{lstlisting}
sudo setcap 'cap_net_bind_service=+ep' /home/pas/.nvm/versions/node/<new-version>/bin/node
\end{lstlisting}
relogon as pas user
\begin{lstlisting}
crontab -e and edit NODE_EXEC line to be /home/pas/.nvm/versions/node/<new-version>/bin/node
npm install -g pm2 eslint eslint-config eslint-plugin-html mocha polymer-cli
pm2 updatePM2
pm2 startup #copy output and run as pi user (who has the sudo capability needed)
pm2 stop all, pm2 delete all, pm2 start pas.json, pm2 save
\end{lstlisting}

\section{Creating a Copy of the SD Card Image}

It is possible to create copies of subvolumes using a pair of commands.  \texttt{btrfs send} can read a subvolume
and output some internal commands to sandard output which can be read on its standard input by \texttt{btrfs receive}.
If running on the same machine the simple unix pipe can be used to connect these two programs together.  Alternatively its
possible to use ssh to pipe the data between machines.  So a possible approach might be to use ssh to copy from the production
system to a desktop machine and then repeat the process from the desktop on to a new sd card.

The paragraphs below leads the reader through this process

\begin{itemize}
  \item Ssh into the production system something like \texttt{ssh pi@pas.accuvision.local}.
  \item make readonly snapshots of all the subvolumes as so:-
  \begin{lstlisting}
    sudo mount -o subvol=/ /dev/mmcblk0p2 /mnt/btrfs_pool
    sudo btrfs subvolume snapshot -r /mnt/btrfs_pool/_root /mnt/btrfs_pool/rootbak
    sudo btrfs subvolume snapshot -r /mnt/btrfs_pool/_log /mnt/btrfs_pool/logbak
    sudo btrfs subvolume snapshot -r /mnt/btrfs_pool/_pas /mnt/btrfs_pool/pasbak
    btrfs subvolume create /mnt/btrfs_pool/bootbak
    rsync -axAHX /boot/ /mnt/btrfs_pool/bootbak/
    btrfs property set -ts /mnt/btrfs_pool/bootbak ro true
  \end{lstlisting}
  \item return to the host system and at some appropriate place to create a subvolume
  \item perform the following
  \begin{lstlisting}
    sudo btrfs subvolume create pas
    ssh pi@pas.accuvision.local "sudo btrfs send /mnt/btrfs_pool/rootbak" | sudo btrfs receive pas/
    ssh pi@pas.accuvision.local "sudo btrfs send /mnt/btrfs_pool/logbak" | sudo btrfs receive pas/
    ssh pi@pas.accuvision.local "sudo btrfs send /mnt/btrfs_pool/pasbak" | sudo btrfs receive pas/
    ssh pi@pas.accuvision.local "sudo btrfs send /mnt/btrfs_pool/bootbak" | sudo btrfs receive pas/
  \end{lstlisting}
  \item return to the production system
  \item perform the following:-
  \begin{lstlisting}
    sudo btrfs subvolume delete -c /mnt/btrfs_pool/rootbak
    sudo btrfs subvolume delete -c /mnt/btrfs_pool/logbak
    sudo btrfs subvolume delete -c /mnt/btrfs_pool/pasbak
    sudo btrfs subvolume delete -c /mnt/btrfs_pool/bootbak
    sudu umount /mnt/btrfs_pool
  \end{lstlisting}
  \item return to host system and place an sd card into a usb adaptor and ports
  \item create partitions on a new 16GB sd card using \texttt{fdisk sdX}.  Ideally, partition boundaries should
  match the write restrictions on the SD cards, which on the SanDisk cards being used for PAS means 4M.  This means the
  partition start sector boundaries (with the previous end sector 1 sector before an the last partition ending at the max) should be (along with partition type):
  \begin{itemize}
    \item Partition 1; 8192 type 0x0c
    \item Partition 2; 540672 type 0x83
    \item Partition 3; 2906064 type 0x82
  \end{itemize}
  \item create filesystems on each partition
  \begin{lstlisting}
    mkfs -t vfat /dev/sdX1
    fatlabel /dev/sdX1 boot
    mkfs -t btrfs -L rootfs /dev/sdX2
    mkswap -L swap /dev/sdX3
  \end{lstlisting}
  \item use \texttt{btrfs send} and \texttt{btrfs receive} to create the correct files on the sd card
  \begin{lstlisting}
    mount /dev/sdX1 /mnt
    rsync -axAHX pas/bootbak/ /mnt/
    umount /dev/sdX1
    mount -o subvol=/ /dev/sdX2 /mnt
    btrfs send pas/rootbak | btrfs receive /mnt/
    mv /mnt/rootbak /mnt/_root
    btrfs send pas/logbak | btrfs receive /mnt/
    mv /mnt/logbak /mnt/_log
    btrfs send pas/pasbak | btrfs receive /mnt/
    mv /mnt/pasbak /mnt/_pas
    umount /mnt
  \end{lstlisting}
  The sd card is now a copy of the live production system and can be carried to whereever it is needed.  It is up to the
  reader to decide whether to keep the subvolumes at pas so the futher sd card copies can be made.
\end{itemize}

\subsection{Update Current System to latest}

\begin{lstlisting}
  sudo apt update
  sudo apt upgrade
  sudo apt full-upgrade
  sudo apt --purge autoremove
  sudo reboot  #ensures we are using the lastest kernel with the initramfs that might have been made
\end{lstlisting}
\subsection{Modify Release Used}
We do an automatic edit to catch all re required files
\begin{lstlisting}
  sudo sed -i /deb/s/stretch/buster/g /etc/apt/sources.list
  sudo sed -i /deb/s/stretch/buster/g /etc/apt/sources.list.d/*.list
\end{lstlisting}

\subsection{Update Package Lists}
Get the lastest package lists imported
\begin{lstlisting}
  sudo apt update
\end{lstlisting}

\subsection{Do the Upgrade}
We are now ready to do the upgrade.
\begin{lstlisting}
  sudo apt upgrade
  sudo apt full-upgrade
\end{lstlisting}

The upgrade process is carried out in two steps: First of all, there is a minimal upgrade to avoid
conflicts and then a complete upgrade. The upgrade may take a while depending on your Internet
connection. A certain amount of attention is required during the update process, as some
configuration files are also updated. Therefore, Raspbian asks several times if you want to replace
certain configuration files with a new one. If you have made changes to individual configuration
files yourself in the past, you should press \textbf{D} (display differences) or \textbf{N} (not replace) depending on
the file, otherwise \textbf{Y} (to replace the old configuration file with a new one).

\subsection{Clean Up Installation}

We remove obsolete/out of date packages
\begin{lstlisting}
  sudo apt --purge autoremove
\end{lstlisting}
Finally reboot to pick up the new system
\begin{lstlisting}
sudo reboot
\end{lstlisting}
\section {The Accuvision Certificate Authority}

An essential requirement for the web server part of PASv5 is the use of a certificate to enable the web server to talk \texttt{https} to
the client machines.  In order to be usable this certificate must be trusted by the browser.  The normal approach for this is that the browser
has a list of public certificate authorities who it trusts to have signed the certificate offered up by the web browser.  An agreement in 2015
means that certicate authorities will only sign certificates for domains that they know the requested has control of.

PASv5 is designed to run inside the clinic on a domain \texttt{accuvision.local} which is not, in the public world, controlled by Accuvision. In
fact the \texttt{.local} part is explicitly reservered for anyone to use how they wish inside their private worlds.  So there is no way a trusted
certificate authority will sign a certificate for \texttt{pas.accuvusion.local}.  The solution to this problem is for Accuvision themselves to set
themselves up as a private certificate authority, sign the certificate for \texttt{pas.accuvusion.local}, and then provide a certificate that describes
themselves to be manually installed in browsers where their certification ability makes sense (i.e. inside the clinic).  PASv5 itself provides a download
link of the certificate authorities certificate.

Previously Alan Chandler acted as the certificate authority and kept all the files associated with that role on his home computer, but the certificate he
set up is expiring at the end of 2020, and the replacement has to survive for a long time - maybe long after he is still providing support.  So the files
and instructions for use have been moved into a directory on the pas server (\texttt{/home/pi/AccuvisionCA}) where they can be used in the longer term.

\end{document}
